Perturbing SU($n$)-symmetric interactions

### 2020-02-07 collective_sun/sun_perturbation.tex
cleaned up appendix on products of multi-body operators

\newcommand{\FS}{\text{FS}}
\newcommand{\EQFS}{=_{\text{FS}}}
\newcommand{\col}{\underline}

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQFS \sum_{X,p} v_{Xp} X_1
  \EQFS \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \sum_p X_p,
\end{align}
are respectively the mean value in $\v v_X$ and the collective version
of $X$.  In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary permutationally symmetric
state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \sum_p \p{\sum_q h_{pq} v_{Xq} - h_p v_{Xp}} X_p
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ (indexed by $\Delta$) is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\m h$ is a matrix of all $h_{pq}$, and
$\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on the
diagonal and zeros everywhere else.  Solving the general eigenvalue
problem in \eqref{eq:cond_1} requires diagonalizing the $N\times N$
matrix $\m h-\diag\v h$.  When the interaction Hamiltonian $H_0$ is
translationally invariant, however, this eigenvalue problem can be
solved analytically; we provide the corresponding solution in Appendix
\ref{sec:trans_inv}.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the permutationally
symmetric manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQFS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - N \col{XY}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
and for any single-spin operator $Z$, we define the collective
operator
\begin{align}
  \col{Z} \equiv \sum_p Z_p.
\end{align}

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-spin
operator $O_{pq}$.  If the two-spin operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q}
  \sp{\sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}
  - \bar h_{pq} w_{Opq}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_p + h_q - 2 h_{pq},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h}
  \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
    \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}
    - \bar h_{pq} \bra{\set{p,q}}}.
  \label{eq:h_super_mat}
\end{align}
Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
discuss this reduction in Appendix \ref{sec:trans_inv}.

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:POQP}.  In the special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k$ are scalar coefficients, and we have neglected scalar
terms in $H_2^{(2)}$ without physical consequence.  In order to expand
the coefficients $W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to concisely write
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

A classification of terms in \eqref{eq:sym_prod_vec} by index
assignments $g\in\G\p{\m M}$ allows us to expand
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
  \label{eq:sym_prod_group_start}
\end{align}
where $\EQFS$ denotes equality up to a restriction to the fully
symmetric manifold, $O\p{g}$ is an operator acquired by assigning
indices to the tensors $O_j$ in a manner consistent with $g$; $w\p{g}$
is a scalar acquired by summing over all indices assigned to the
tensors $w_j$ according to $g$; and $\S\p{g}$ is a symmetry factor
accounting for the number of equivalent ways to assign indices
according to $g$.

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we identify the set of values that the distinct indices
assigned according to $g$ can take:
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} },
  \label{eq:index_values}
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
$\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
the tensors $\set{w_j}$, a set $k_S$ of $g_S$ distinct spins, with the
restriction that all $k_S$ are orthogonal: if $z\in k_S$, then
$z\notin k_R$ for $R\ne S$.  Having defined the set $\C_N\p{g}$ of
index values, we define the restriction of $k\in\C_N\p{g}$ to values
associated with $w_j$ and $O_j$:
\begin{align}
  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.
These definitions allow us to expand
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  O\p{g} \equiv \prod_{j\in\ZZ_p}
  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
  \label{eq:diagram_factors}
\end{align}
where the choice of $\ell\in\C_N\p{g}$ does not matter after
projection onto permutationally symmetric manifold, and
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains
multinomial coefficients
\begin{align}
  { M \choose a_1, a_2, \cdots, a_m }
  \equiv \f{M!}{a_1!a_2!\cdots a_m!}
\end{align}
that account for the number of ways to partition the indices of each
$w_j,O_j$ into sets of shared indices according to $g_S$ for all
$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
accounts for the number of ways to permute each set of shared indices
on every tensor.  Altogether, the symmetry factors $\S\p{g}$ cancel
out with the $\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
  \label{eq:sym_prod_group}
\end{align}

@@

Diagrams such as \eqref{eq:venn_diagram} will play a major role in the
calculations in this section.  Here, we first define diagrams as
formal objects that we can use in mathematical expressions.
Formalizing diagrams will enable us to perform necessary calculations
that are otherwise intractable.  We then discuss rules for decomposing
these diagrams in order to reduce the computational cost of their
numerical evaluation.

The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
intersection of the circles $j\in S$, and are simply equal to the
associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
\begin{align}
  \diagram{example_123}
  \equiv \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
      \abs{\set{a,b,c,d,e}}=5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
\end{align}
Such a diagram with $\abs{g}$ dots nominally takes $\O\p{N^{\abs{g}}}$
time to compute.  In practice, this computational cost can be
prohibitive for performing numerical simulations of systems with any
appreciable size.  We therefore wish to simplify diagrams in order to
reduce their computational cost as much as possible.  Our general
strategy will be to replace constrained sums that make all indices
interdependent (such as $\abs{\set{a,b,c,d,e}}=5$ above) by
unconstrained sums that can be carried out in sequence.  In order to
represent different constraints, we need to define diagrams in which
filled dots (``$\bullet$'') may be replaced by empty dots
(``$\circ$'') or crosses (``$\bm\times$'').

>>>>

\newcommand{\PS}{\text{PS}}
\newcommand{\EQPS}{=_{\text{PS}}}
\newcommand{\col}{\underline}

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQPS \sum_{X,p} v_{Xp} X_1
  \EQPS \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \sum_p X_p,
\end{align}
are respectively the mean value in $\v v_X$ and the collective version
of $X$.  In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary permutationally symmetric
state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \sum_p \p{\sum_q h_{pq} v_{Xq} - h_p v_{Xp}} X_p
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ (indexed by $\Delta$) is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\m h$ is a matrix of all $h_{pq}$, and
$\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on the
diagonal and zeros everywhere else.  Solving the general eigenvalue
problem in \eqref{eq:cond_1} requires diagonalizing the $N\times N$
matrix $\m h-\diag\v h$.  When the interaction Hamiltonian $H_0$ is
translationally invariant, however, this eigenvalue problem can be
solved analytically; we provide the corresponding solution in Appendix
\ref{sec:trans_inv}.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the permutationally
symmetric manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQPS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - N \col{XY}},
  \label{eq:H_1_2}
\end{align}
where $\EQPS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
and for any single-spin operator $Z$, we define the collective
operator
\begin{align}
  \col{Z} \equiv \sum_p Z_p.
\end{align}

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQPS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
  \EQPS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-spin
operator $O_{pq}$.  If the two-spin operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQPS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q}
  \sp{\sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}
  - \bar h_{pq} w_{Opq}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_p + h_q - 2 h_{pq},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h}
  \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
    \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}
    - \bar h_{pq} \bra{\set{p,q}}}.
  \label{eq:h_super_mat}
\end{align}
Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
discuss this reduction in Appendix \ref{sec:trans_inv}.

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:POQP}.  In the special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \simeq_\PS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $\simeq_\PS$ denotes equality up to scalar terms and a
restriction to the permutationally symmetric manifold;
$S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin operator; and
$W_k$ are scalar coefficients.  In order to expand the coefficients
$W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to concisely write
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

A classification of terms in \eqref{eq:sym_prod_vec} by index
assignments $g\in\G\p{\m M}$ allows us to expand
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
  \EQPS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
  \label{eq:sym_prod_group_start}
\end{align}
where $\EQPS$ denotes equality up to a restriction to the fully
symmetric manifold, $O\p{g}$ is an operator acquired by assigning
indices to the tensors $O_j$ in a manner consistent with $g$; $w\p{g}$
is a scalar acquired by summing over all indices assigned to the
tensors $w_j$ according to $g$; and $\S\p{g}$ is a symmetry factor
accounting for the number of equivalent ways to assign indices
according to $g$.

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we identify the set of values that the distinct indices
assigned according to $g$ can take:
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} },
  \label{eq:index_values}
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
$\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
the tensors $\set{w_j}$, a set $k_S$ of $g_S$ distinct spins, with the
restriction that all $k_S$ are orthogonal: if $z\in k_S$, then
$z\notin k_R$ for $R\ne S$.  Having defined the set $\C_N\p{g}$ of
index values, we define the restriction of $k\in\C_N\p{g}$ to values
associated with $w_j$ and $O_j$:
\begin{align}
  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.
These definitions allow us to expand
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  O\p{g} \equiv \prod_{j\in\ZZ_p}
  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
  \label{eq:diagram_factors}
\end{align}
where the choice of $\ell\in\C_N\p{g}$ does not matter after
projection onto permutationally symmetric manifold, and
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains
multinomial coefficients
\begin{align}
  { M \choose a_1, a_2, \cdots, a_m }
  \equiv \f{M!}{a_1!a_2!\cdots a_m!}
\end{align}
that account for the number of ways to partition the indices of each
$w_j,O_j$ into sets of shared indices as specified by the values of
$g_S$ for all $j\in S$, and the second product in
\eqref{eq:sym_prod_symmetry} accounts for the number of ways to
permute each set of shared indices on every tensor.  Altogether, the
symmetry factors $\S\p{g}$ cancel out with the $\sim M_j!$ prefactors
in \eqref{eq:sym_prod_vec}, so
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
  \label{eq:sym_prod_group}
\end{align}

@@

Diagrams such as \eqref{eq:venn_diagram} will play a major role in the
calculations in this section.  Here, we define these diagrams as
formal objects that we can use in mathematical expressions, enabling
us to systematically perform calculations that are otherwise
intractable.  We then derive rules to ``simplify'' these diagrams in
such a way as to reduce the computational cost of their numerical
evaluation.

The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
intersection of the circles $j\in S\in\PP\p{\ZZ_p}$, and are simply
equal to the associated scalar $w\p{g}$ defined in
\eqref{eq:diagram_factors}, e.g.
\begin{align}
  \diagram{example_123}
  \equiv \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
      \abs{\set{a,b,c,d,e}}=5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
\end{align}
Such a diagram with $\abs{g}$ dots nominally takes $\O\p{N^{\abs{g}}}$
time to compute.  In practice, this computational cost can be
prohibitive for performing numerical simulations of systems with any
appreciable size.  We therefore wish to simplify diagrams in order to
reduce their computational cost as much as possible.  Our general
strategy will be to replace constrained sums that make all indices
interdependent (e.g.~$\abs{\set{a,b,c,d,e}}=5$ above) by unconstrained
sums that can be carried out in sequence.  In order to represent
different constraints, we need to define diagrams in which filled dots
(``$\bullet$'') may be replaced by empty dots (``$\circ$'') or crosses
(``$\bm\times$'').

If a region of a diagram contains $F$ filled dots, those dots
correspond to indices that sum over $\C_N\p{F}$ with the constraint
that they are not equal to any other filled dots in a diagram.  $E$
empty dots in a region, meanwhile, correspond to indices that are
summed over $\C_N\p{E}$ without any additional constraints, e.g.
\begin{align}
  \diagram{example_o}
  \equiv \sum_{\substack{\p{a,b}\in\C_N\p{2}\\e\in\ZZ_N\\
      \abs{\set{a,b,e}}=3}}
  \sum_{\substack{\p{c,d}\in\C_N\p{2}\\f\in\ZZ_N}}
  v\p{a,b,c,d,e} w\p{e,f}.
\end{align}
Empty dots are therefore entirely independent of filled dots.
Crosses, meanwhile, are {\it entirely dependent} on filled dots:
``cross indices'' are summed {\it only} over the values of ``filled
dot indices'', e.g.
\begin{align}
  \diagram{example_x}
  \equiv \sum_{\substack{a,e\in\ZZ_N\\\abs{\set{a,e}}=2}}
  \sum_{b,c\in\set{a,e}} \sum_{d\in\ZZ_N}
  u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
\end{align}
Empty dots and crosses allow us to decompose diagrams with a high
computational cost into different diagrams with a lower computational
cost.  We will essentially use two tricks to decompose diagrams:
eliminating filled dots (which will add empty dots and crosses), and
eliminating crosses (which will move around filled dots).
